Auto-tuned Jacobi-like Multi-grid Smoother\\ for Fast Pressure Projection

### 2021-10-01 tex/article.tex

However, none of these studies have applied data-driven techniques to accelerate the critical projection step itself. Some recent work has focused on the pressure projection step explicitly. Chaotic projection methods for massively parallel systems; not data driven. Multi-grid data-driven decomposition; high errors and unclear speed-up. This work links develops a high-speed parameterized smoother which outperforms classic smoothing methods on eleven 2D and 3D benchmarks. Moreover, as the approach in linear in the pressure and scale-invariant, the parameters can be tuned on a coarse simulation and maintain their accelerated performance on the fine simulation.

@@

The linearity of equation \ref{eq:aer} is a critical property which the data-driven method must maintain in order to be useful in practise, but this linearity is only with respect to the update and residual fields. The operators of the ``MG network'' can all potentially be optimized for a specific sub-class of problems. In particular, those operators can be made into parameterized nonlinear functions of $A$, embedding information about the discrete problem and boundary conditions, without violating linearity.

@@

There are few constraints on $\tilde A^{-1}$: it must have units of $A^{-1}$, and symmetry of $A$ implies the approximate inverse should also be symmetric. For simplicity, the diagonal and off-diagonal coefficients of $\tilde A^{-1}$ are constructed independently 
\begin{equation}
    \tilde a^{-1}_{ii} = \frac{f_d(a_{ii}/s\,|\theta_d)}{a_{ii}} , \quad
    \tilde a^{-1}_{ij} = \frac{f_o(a_{ij}/s\,|\theta_o)}{a_{ii}+a_{jj}}
\end{equation}
where function inputs are scaled by the maximum off-diagonal $s=\max(A-D)$ and the outputs are scaled by the diagonal elements. Note a Jacobi smoother would use a diagonal function $f_d=1$ and off-diagonal $f_o=0$.

Automatic differentiation (AD) is used to optimize the parameterized smoother for a set of training data $X=\{A,x^k,b\}$. The loss function is simply the reduction in the norm of the residual after a V-cycle iteration $L = \log_{10}(|r^{k+1}| / |r^{k}|)$, which means that example pressure \textit{solutions} are not required. This loss is averaged over each iteration and each example in the data-set. The optimal parameters are defined as
\begin{equation}
    \hat\theta = \min_\theta L(\theta\, |X)
\end{equation}
and determined using Newton's method. The ForwardDiff.jl AD package is used determine the gradient $\nabla_\theta L$ and Hessian $H_{\theta\theta} L$ automatically. The adaptive and recursive nature of the MG V-cycle and general nonlinear form of $f$ would make analytic derivatives prohibitively unwieldy and error prone, and numerical finite differences potentially unstable and inaccurate.

@@

A set of six synthetic projection cases were constructed to establish the characteristics of the parameterized smoother. The three 2D cases are show in Figure~\ref{fig:2D-box}-\ref{fig:2D-sphere}, and each has a matching 3D case. The `static' cases have a simple hydro-static gradient solution throughout the domain, with the direction and magnitude randomized. The `dipole' cases have a localized dipole within a quiescent flow with randomized placement, direction and magnitude. The `sphere' cases have a moving sphere immersed within a quiescent flow with randomized radius, location, and velocity direction. Neumann boundary conditions are applied on all boundaries, including the immersed sphere, through adjustment of the $A$ matrix coefficients which are otherwise uniform. All cases are initialized with $x^0=0$, and all cases feature initially localized residuals; i.e. $r^0\ne0$ only in a fraction of the cells, but $x$ must be updated throughout the domain, starkly illustrating the elliptic nature of the projection step.

Figure~\ref{fig:cross plot} shows the `test' performance of the parameterized smoother over a single V-cycle on 100 randomized examples after optimization using 100 examples of the `test' case. Each example uses a grid size of n=32 point in each direction ($N=32^2$ in 2D and $N=32^3$ in 3D) as testing with other resolutions found essentially no change in the performance. As expected, the performance is best when testing and training on the same case, with the residual reduced by $10^{-1.18}$ on the 2D-sphere case down to $10^{-1.75}$ on the 2D-box case. While the performance for the simplest cases generalizes poorly, the 2D-sphere smoother generalizes essentially as well as the `union' smoother trained on all of the data. 
%The performance of the `union' smoother is only 1-4\% lower than the smoother trained on the test case for the dipole and sphere cases.

Figure~\ref{fig:synthetic time} characterizes the acceleration achieved by the data-driven smoother using the time taken to reduce the residual of each case by $10^{-3}$, scaled by the time to run a single V-cycle using a Jacobi smoother. The results for the Jacobi smoother are not given, since thousands of iterations are required. Remarkably, the Jacobi-like parameterized smoother is only 36\% slower per iteration, and yet converges in only a few V-cycles. This results in a 80-175\% speed up relative to optimized serial Gauss-Sidel and SOR smoothers, and the speed-up will be even more favorable in parallel computation.

>>>>

However, none of these studies have applied data-driven techniques to accelerate the critical projection step itself. Some recent work has focused on the pressure projection step explicitly. Chaotic projection methods for massively parallel systems; not data driven. Multi-grid data-driven decomposition; high errors and unclear speed-up. This work develops a high-speed data-driven project method which outperforms classic methods on eleven 2D and 3D benchmarks. Moreover, as the approach in linear in the pressure and scale-invariant, the parameters can be tuned on a coarse simulation and maintain their accelerated performance on unseen fine-grid simulations.

@@

The linearity of equation \ref{eq:aer} is a critical property which the data-driven method must maintain in order to be useful in practise, but this linearity is only with respect to the update and residual fields. The operators of the ``MG network'' can all potentially be optimized for a specific sub-class of problems. In particular, those operators can be made into parameterized nonlinear functions of $A$, embedding information about the discrete problem and boundary conditions.

@@

There are few constraints on $\tilde A^{-1}$: it must have units of $A^{-1}$, and symmetry of $A$ implies the approximate inverse should also be symmetric. For simplicity, the diagonal and off-diagonal coefficients of $\tilde A^{-1}$ are constructed independently 
\begin{equation}
    \tilde a^{-1}_{ii} = \frac{f_d(a_{ii}/s\,|\theta_d)}{a_{ii}} , \quad
    \tilde a^{-1}_{ij} = \frac{f_o(a_{ij}/s\,|\theta_o)}{a_{ii}+a_{jj}}
\end{equation}
where function inputs are scaled by the maximum off-diagonal $s=\max(A-D)$ and the outputs are scaled by the diagonal elements. Note a Jacobi smoother would use a diagonal function $f_d=1$ and off-diagonal function $f_o=0$. With the learning problem now reduced to two scaled single-variable functions, the specific choice of parameterization is not critical and a simple quadratic polynomial $f(a) = \sum_{p=0}^2 a^p \theta_p$ is chosen for $f_o$ and $f_d$. Using higher-order polynomials, splines, and interpolating kernels did not significantly change the results.

Automatic differentiation (AD) is used to optimize the parameterized smoother for a set of training data $X=\{A,x^k,b\}$. The loss function is simply the reduction in the norm of the residual after a V-cycle iteration $L = \log_{10}(|r^{k+1}| / |r^{k}|)$, which means that example pressure \textit{solutions} are not required. This loss is averaged over each iteration and each example in the data-set. The optimal parameters are defined as
\begin{equation}
    \hat\theta = \min_\theta L(\theta\, |X)
\end{equation}
and determined using Newton's method. The ForwardDiff.jl AD package is used determine the gradient $\nabla_\theta L$ and Hessian $H_{\theta\theta} L$ automatically, avoiding inaccurate and potentially unstable finite differences of the nonlinear recursive loss function.

@@

A set of six synthetic projection cases were constructed to establish the characteristics of the parameterized smoother. The three 2D cases are show in Figure~\ref{fig:2D-box}-\ref{fig:2D-sphere}, and each has a matching 3D case. The `static' cases have a hydrostatic gradient throughout the domain, with the direction and magnitude randomized. The `dipole' cases have a localized dipole within a quiescent flow with randomized placement, direction and magnitude. The `sphere' cases have a moving sphere immersed within a quiescent flow with randomized radius, location, and velocity direction. All cases are initialized with $x^0=0$.
%, and all cases feature initially localized residuals; i.e. $r^0\ne0$ only in a fraction of the cells, but $x$ must be updated throughout the domain, starkly illustrating the elliptic nature of the projection step.
Neumann boundary conditions are applied on all boundaries, including the immersed sphere, through adjustment of the $A$ matrix coefficients which are otherwise uniform.

Figure~\ref{fig:cross plot} characterizes the generalization of the parameterized smoother's residual reduction over a single V-cycle on 100 randomized `test' case examples after optimization using 100 examples of the `train' case. Each example uses a grid size of $n=32^M$ points ($N=32^2$ in 2D and $N=32^3$ in 3D) as testing with other resolutions found essentially no change in the performance. As expected, the performance is best when testing and training on the same case, with a single V-cycle reducing the residual from $10^{-1.17}$ on the 2D-sphere case down to $10^{-1.58}$ on the 3D-box case. While the performance for the simplest cases generalizes poorly, the 2D-sphere smoother generalizes essentially as well as the `union' smoother trained on all of the data. 
%The performance of the `union' smoother is only 1-4\% lower than the smoother trained on the test case for the dipole and sphere cases.

Finally, the acceleration of the `union' data-driven smoother is evaluated on new example data of a different size, $n=64^M$ points. Figure~\ref{fig:synthetic time} shows the time to reduce the residual of each case by $10^{-3}$ relative to the time to run a single V-cycle using a Jacobi smoother, but the results for the Jacobi smoother are not shown since hundreds or even thousands of V-cycles are required to converge. Remarkably, the Jacobi-like parameterized smoother is only around 33\% slower per V-cycle, and yet converges in only 1-3 V-cycles in all cases. This results in a 80-175\% speed up relative to optimized serial Gauss-Sidel and SOR smoothers. This speed-up will be even more favorable in parallel computation as the new smoother operates directly on the local residual without back-propagation, eliminating any need for additional communication.
